### 2023-07-05 tesis_Ashish/Abstract/abstract.tex
AS thesis update

The Pixel Cluster Counting (PCC) luminosity measurement method has emerged as a reliable offline technique for luminosity measurement during LHC run periods. The PCC method is based on the rate of pixel clusters in the CMS pixel detector, which is the innermost part of the CMS tracker. PCC is studied for its stability and accuracy in luminosity measurements using 2018 AlCa Reco datasets. The results of these studies demonstrate that the PCC luminosity measurement method is stable and accurate, able to achieve ??\% systematic uncertainty on the final luminosity measurement for Run 2 data. Moreover, the algorithms and optimization for the Phase II High Luminosity LHC (HL-LHC) luminosity measurement are also developed, using the Tracker Endcap Pixel Detector (TEPX). This detector is expected to provide excellent statistical and linearity performance, which is essential for accurate luminosity measurements at the HL-LHC. The TEPX detector will replace the upgraded Phase I pixel detector and is expected to provide even higher precision luminosity measurement.

>>>>

The Pixel Cluster Counting (PCC) luminosity measurement method has emerged as a reliable offline technique for luminosity measurement.
%The PCC method is based on the rate of pixel clusters in the CMS pixel detector, which is the innermost part of the CMS tracker.
PCC method is studied for its precision in luminosity measurement using 2018 AlCa Reco datasets. The results of these studies demonstrate that the PCC luminosity measurement method is stable and linear, able to achieve 1\% systematic uncertainty on the final luminosity measurement. Moreover, the clusters and coincidences counting algorithms and their optimization for the Phase II High Luminosity LHC (HL-LHC) luminosity measurement are also developed, using simulated datasets for the proposed Phase II pixel detector. %This new pixel detector will replace the Run 2 pixel detector.
Linearity of the TEPX and D4R1 luminometers are expected to be within 1\% (offline) up to a pileup of 200 and their statistical precision per second is about 0.1\% at <PU>=200.
%It is expected to provide excellent statistical precision and linearity performance, which is essential for precision luminosity measurements at the HL-LHC.

### 2023-07-05 tesis_Ashish/Chapter1/chapter1.tex
new thesis
AS thesis update

Collider experiments are a type of particle physics experiment that involve colliding particles, typically protons or electrons, at very high energies. These collisions can lead to the creation of new particles or the detection of new phenomena, providing insight into the fundamental nature of matter. Colliders are designed to accelerate particles to nearly the speed of light and guide them along a circular or linear path using magnetic fields. By colliding particles head-on or at oblique angles, physicists can study the resulting particle showers and detect the products of the collision using complex detectors. Collider experiments have played a crucial role in advancing our understanding of particle physics and continue to be a vital tool for scientific exploration.

@@

Luminosity is a measure of the number of particle collisions that occur per unit time and per unit area. In particle physics, luminosity is a crucial parameter that determines the rate at which particles are produced in a collider experiment.

@@

The importance of luminosity in particle physics lies in its direct relationship with the production rate of particles of interest. The production rate of a given particle is proportional to the product of the luminosity and the cross-section of the process. Thus, luminosity directly affects the ability to detect rare events and make precise measurements. The more collisions that occur, the higher the probability of rare events happening, such as the creation of new particles or the detection of new phenomena. Additionally, precise measurements require a large number of collisions to be able to statistically determine the properties of the particles being studied.

@@

where R is the production rate, L is the luminosity, and $\sigma$ is the cross-section of the process that produces the particle. The cross-section is a measure of the probability that a given process will occur, and it is typically a very small number, on the order of picobarns ($10^{-36} m^2$).
In order to achieve high luminosity, particle accelerators use various techniques to increase the number of particles in the beams and focus them to a small cross-sectional area. This is done by using superconducting magnets to guide the particles along a circular path and keep them confined to a small area. Another technique is to use radio frequency cavities to accelerate the particles to higher energies increasing the frequency of bunch crossings. Increasing the number of particles in each bunch is another way to increase luminosity. In high-energy particle accelerators, such as the Large Hadron Collider (LHC) at CERN, the luminosity is typically measured in units of inverse femtobarns ($fb^{-1}$), which is a measure of the total number of collisions per unit area during the course of an experiment. For example, the LHC has a design luminosity of $10^{34} cm^{-2}s^{-1}$, which means that there are $10^{34}$ proton-proton collisions per second per square centimeter of the detector. The importance of luminosity can be seen in the many discoveries that have been made in particle physics due to high-luminosity experiments. For example, the discovery of the Higgs boson at the LHC was made possible by the high luminosity of the collider, which allowed physicists to observe the rare Higgs boson signal over the background noise of other particles. In addition to discoveries, high luminosity experiments are also important for precision measurements. For example, the measurement of the properties of the top quark, the heaviest known elementary particle, requires a large number of collisions to be able to distinguish the signal from the background noise.

@@

In particle physics, the luminosity of colliding beams is a crucial measurement as it provides information about the rate at which rare processes occur and enables precise measurements of particle properties such as cross sections and couplings to fermions, bosons. Cross sections are fundamental quantities in particle physics, used to predict the number of events a certain process will produce. Luminosity depends on several variables, including the number of protons in each beam, the size and shape of the beams, and their relative positioning. A slight change in any of these variables can significantly alter the luminosity, which in turn affects the interpretation of collision data. Therefore, it is crucial to measure the luminosity with high precision.
%The measurement of luminosity with the pixel detector at the CMS experiment is particularly important  as it is a highly precise detector that is capable of measuring luminosity with great accuracy. It provides information about the number of charged particles passing through the detector, which is directly proportional to the number of collisions. This information is then used to calculate the luminosity. 
Improving the uncertainty in luminosity measurements is essential for many physics analyses, including searches for new particles and measurements of the properties of known particles. It is a dominant factor in the measurement of Higgs Boson cross sections and couplings to other gauge bosons and fermions, as shown in Fig~\ref{fig:lum_unc}. As such, the uncertainty in cross-section and coupling measurements cannot be better than the uncertainty in luminosity measurement. Therefore, ongoing efforts are focused on improving the accuracy of luminosity measurements, with the pixel detector at CMS being a key component of this effort. By reducing the luminosity uncertainty, physicists can make more precise measurements of particle properties and potentially discover new physics beyond the Standard Model.

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.42\textwidth]{ashish_thesis/lumi_precision.png}
\caption{%
   Higgs boson production processes cross section projected uncertainty. In the most precisely measured Higgs boson production process, gluon fusion (ggH), luminosity uncertainty will dominate the experimental systematic uncertainty at the High Luminosity (HL)-LHC.
}
\label{fig:lum_unc}
\end{figure}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.6\textwidth]{ashish_thesis/lhc_schematic.png}
\caption{%
    CERN accelerator complex consists of Linac 2, Proton Synchrotron Booster (PSB), Proton Synchrotron (PS), Super Proton Synchrotron (SPS) and Large Hadron Collider (LHC). LHC ring showing four interaction points where the ALICE, ATLAS, LHCb and CMS detectors are stationed.
}
\label{fig:lhc}
\end{figure}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.7\textwidth]{ashish_thesis/cms_exp.png}
\caption{%
   Compact muon solenoid (CMS) detector with all of the sub detectors showing silicon tracker, electromagnetic and hadron calorimeters and muon chambers. 
}
\label{fig:cms}
\end{figure}

@@

During the operation of the CMS experiment, two main data-taking periods were conducted: Run 1, which spanned from 2010 to 2013, and Run 2, which took place from 2015 to 2018. During these periods, the CMS experiment measured the instantaneous and integrated luminosity of the proton-proton collisions produced by the Large Hadron Collider (LHC) shown in Fig.~\ref{fig:lumi}. During Run 1, the CMS experiment achieved a peak instantaneous luminosity of up to $5 \times 10^{33} cm^{-2} s^{-1}$, while during Run 2, the luminosity reached a peak value of up to $2 \times 10^{34} cm^{-2} s^{-1}$. The luminosity measurements during both runs were accompanied by careful assessments of the systematic uncertainties to ensure accurate and reliable results as shown in Table. \ref{tab:lumi}.

@@

\begin{table}[h]
\centering
\caption{Summary of center of mass energy, integrated luminosity, and systematic uncertainty per year during the CMS luminosity measurement.}
\begin{tabular}{c c c c}
Year & Center of Mass Energy (TeV) & Integrated Luminosity (fb$^{-1}$) & Systematic Uncertainty (\%) \\
\hline
2010 & 7 & 0.045 & 11 \\
2011 & 7 and 8 & 6.1 & 4.5 \\
2012 & 8 & 23.3 & 2.5 \\
2015 & 13 & 4.3 & 2.3 \\
2016 & 13 & 41.6 & 2.5 \\
2017 & 13 & 49.8 & 2.3 \\
2018 & 13 & 67.9 &  2.5\\
\end{tabular}
\label{tab:lumi}
\end{table}

\begin{figure}[!htp]
\centering
\includegraphics[width=0.8\textwidth]{ashish_thesis/CMS_luminosity.png}
\caption{%
    Delivered luminosity versus time for Run 1 and Run 2 of the CMS experiment.
}
\label{fig:lumi}
\end{figure}

>>>>

Collider experiments are a type of particle physics experiment that involve colliding particles, typically protons or electrons, at very high energies. These collisions can lead to the creation of new particles or the detection of new phenomena, providing insight into the fundamental nature of matter. Colliders are designed to accelerate particles to nearly the speed of light and guide them along a circular or linear path using magnetic fields. By colliding particles head-on or at oblique angles, physicists can study the resulting particle showers and detect the products of the collision using complex detectors. Collider experiments have played a crucial role in advancing our understanding of particle physics and continue to be a vital tool for scientific exploration. Two important parameters in the collider experiments are energy and luminosity.

@@

Luminosity is a measure of the number of particle collisions that occur per unit area per unit area. In particle physics, luminosity is a crucial parameter that determines the rate at which particles are produced in a collider experiment \cite{Herr:941318}.

@@

The importance of luminosity in particle physics lies in its direct relationship with the production rate of particles of interest. The production rate of a given particle is the product of the luminosity and the cross-section of the process. Thus, luminosity directly affects the ability to detect rare events and make precise measurements. The more collisions that occur, the higher the probability of rare events happening, such as the creation of new particles or the detection of new phenomena. Additionally, precise measurements require a large number of collisions to be able to statistically determine the properties of the particles being studied.

@@

where R is the production rate, L is the luminosity, and $\sigma$ is the cross-section of the process that produces the particle. The cross-section is a measure of the probability that a given process will occur, and it is typically a very small number, on the order of picobarns ($10^{-36} m^2$).
In order to achieve high luminosity, particle accelerators use various techniques to increase the number of particles in the beams and focus them to a small cross-sectional area. This is done by using superconducting magnets to guide the particles along a circular path and keep them confined to a small area. Another technique is to use radio frequency cavities to accelerate the particles to higher energies increasing the frequency of bunch crossings. In high-energy particle accelerators, such as the Large Hadron Collider (LHC) at CERN, the luminosity is typically measured in units of inverse femtobarns ($fb^{-1}$), which is a measure of the total number of collisions per unit area during the course of an experiment. For example, the LHC has a design luminosity of $10^{34} cm^{-2}s^{-1}$, which means that there are $10^{34}$ proton-proton collisions per square centimeter per second. The importance of luminosity can be seen in the many discoveries that have been made in particle physics due to high luminosity experiments. For example, the discovery of the Higgs boson at the LHC was made possible by the high luminosity of the collider, which allowed physicists to observe the rare Higgs boson signal over the background noise of other particles. In addition to discoveries, high luminosity experiments are also important for precision measurements. For example, the measurement of the properties of the top quark, the heaviest known elementary particle, requires a large number of collisions to be able to distinguish the signal from the background noise.

@@

In particle physics, the luminosity of colliding beams is a crucial measurement as it provides information about the rate at which rare processes occur and enables precise measurements of particle properties such as cross sections and couplings to fermions, bosons. Luminosity depends on several variables, including the number of protons in each beam, the size and shape of the beams, and their relative positioning. A slight change in any of these variables can significantly alter the luminosity, which in turn affects the interpretation of collision data. Therefore, it is crucial to measure the luminosity with high precision.
%The measurement of luminosity with the pixel detector at the CMS experiment is particularly important  as it is a highly precise detector that is capable of measuring luminosity with great accuracy. It provides information about the number of charged particles passing through the detector, which is directly proportional to the number of collisions. This information is then used to calculate the luminosity. 
Improving the uncertainty in luminosity measurements is essential for many physics analyses, including searches for new particles and measurements of the properties of known particles. It is a dominant factor in the measurement of Higgs Boson cross sections and couplings to other gauge bosons and fermions, as shown in Fig~\ref{fig:lum_unc}. As such, the uncertainty in cross-section and coupling measurements cannot be better than the uncertainty in luminosity measurement. Therefore, ongoing efforts are focused on improving the precision of luminosity measurements, with the pixel detector at CMS being a key component of this effort. By reducing the luminosity uncertainty, physicists can make more precise measurements of particle properties and potentially discover new physics beyond the Standard Model.

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.8\textwidth]{ashish_thesis/lumi_precision.png}
\caption[Phase II uncertainty projections]{ %
   Higgs boson production processes cross section projected uncertainty. In the most precisely measured Higgs boson production process, gluon fusion (ggH), luminosity uncertainty will dominate other experimental systematic uncertainty at the High Luminosity (HL)-LHC. \cite{Dainese:2703572}
}
\label{fig:lum_unc}
\end{figure}

%\begin{center}
 % \begin{figure}[h]
  %  \centering
   % \includegraphics[scale=.35]{ashish_thesis/lumi_precision.png}
    %\caption[Phase II uncertainty projections]{Higgs boson production processes cross section projected uncertainty. In the most precisely measured Higgs boson production process, gluon fusion (ggH), luminosity uncertainty will dominate the experimental systematic uncertainty at the High Luminosity (HL)-LHC.}
    %\label{fig:lum_unc}
  %\end{figure}
%\end{center}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.9\textwidth]{ashish_thesis/lhc_schematic.png}
\caption[CERN accelerator complex]{%
    CERN accelerator complex consists of Linac 2, Proton Synchrotron Booster (PSB), Proton Synchrotron (PS), Super Proton Synchrotron (SPS) and Large Hadron Collider (LHC). LHC ring showing four interaction points where the ALICE, ATLAS, LHCb and CMS detectors are stationed \cite{lhc_complex}.
}
\label{fig:lhc}
\end{figure}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.75\textwidth]{ashish_thesis/cms_exp.png}
\caption[CMS Experiment]{%
   Compact muon solenoid (CMS) detector with all of the sub detectors showing silicon tracker, electromagnetic and hadron calorimeters and muon chambers \cite{CMS:2008xjf}. 
}
\label{fig:cms}
\end{figure}

@@

During the operation of the CMS experiment, two main data-taking periods were conducted: Run 1, which spanned from 2010 to 2013, and Run 2, which took place from 2015 to 2018. During these periods, the CMS experiment measured the instantaneous and integrated luminosity of the proton-proton collisions produced by the Large Hadron Collider (LHC) where the integrated luminosity for each year is shown in Fig.~\ref{fig:lumi}. During Run 1, the CMS experiment achieved a peak instantaneous luminosity of up to $5 \times 10^{33} cm^{-2} s^{-1}$, while during Run 2, the luminosity reached a peak value of up to $2 \times 10^{34} cm^{-2} s^{-1}$. The luminosity measurements during both runs were accompanied by careful assessments of the systematic uncertainties to ensure accurate and reliable results \cite{CMS-PAS-EWK-10-004}\cite{CMS-PAS-EWK-11-001} \cite{CMS-PAS-SMP-12-008} \cite{CMS-PAS-LUM-13-001} \cite{Sirunyan:2759951} \cite{pas_17}\cite{pas_18} as shown in Table. \ref{tab:lumi}.

@@

\begin{table}[h]
\centering
\caption[Run 1, Run 2 Integrated luminosity and its precision]{Summary of center of mass energy, integrated luminosity, and systematic uncertainty per year during the CMS luminosity measurement.}
\begin{tabular}{c c c c}
Year & Center of Mass Energy (TeV) & Integrated Luminosity (fb$^{-1}$) & Systematic Uncertainty (\%) \\
\hline
2010 & 7 & 0.045 & 11 \\
2011 & 7 and 8 & 6.1 & 2.2 \\
2012 & 8 & 23.3 & 2.5 \\
2015 & 13 & 4.3 & 1.6 \\
2016 & 13 & 41.6 & 1.2 \\
2017 & 13 & 49.8 & 2.3 \\
2018 & 13 & 67.9 &  2.5\\
\end{tabular}
\label{tab:lumi}
\end{table}

\begin{figure}[!htp]
\centering
\includegraphics[width=0.9\textwidth]{ashish_thesis/CMS_luminosity.png}
\caption[Integrated luminosity per year]{%
    Delivered luminosity versus time for Run 1, Run 2 and Run 3 of the CMS Experiment \cite{wikicern_cc}.
}
\label{fig:lumi}
\end{figure}
